---
title: "506HW3"
format: html
author: Romeo Ruan
---


# Problem 1

For the problem, I have used the help frpm ChatGPT for writing code generating logistic regression, the psudo R^2, and AIC values.

## A
```{r}

# Install necessary packages
if (!require(haven)) install.packages("haven")
if (!require(dplyr)) install.packages("dplyr")

library(haven)
library(dplyr)

# Import the datasets from the URLs
vix_d <- read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/VIX_D.XPT")
demo_d <- read_xpt("https://wwwn.cdc.gov/Nchs/Nhanes/2005-2006/DEMO_D.XPT")
# Merge the two datasets using SEQN
merged_data <- merge(vix_d, demo_d, by = "SEQN")
# Keep only matched records
merged_data <- merged_data %>% filter(!is.na(SEQN))
# Check the sample size
nrow(merged_data)
```
## B
```{r}
# Rename variables
merged_data <- merged_data %>%
  rename(glasses = VIQ220, age = RIDAGEYR)
# Handle missing values: Replace glasses = 9 with NA, subtract 1 from glasses variable
merged_data <- merged_data %>%
  mutate(glasses = ifelse(glasses == 9, NA, glasses - 1))
# Create 10-year age brackets
merged_data <- merged_data %>%
  mutate(age_bracket = cut(age, breaks = seq(0, 90, by = 10), include.lowest = TRUE, right = FALSE, 
                           labels = c("0-9", "10-19", "20-29", "30-39", "40-49", "50-59", "60-69", "70-79", "80-89")))
# Calculate the proportion of glasses/contact lenses wearers by age bracket
proportion_glasses <- merged_data %>%
  group_by(age_bracket) %>%
  summarise(Proportion_Wear_Glasses = mean(glasses == 1, na.rm = TRUE) * 100)
# Produce a nicely formatted table with the results
knitr::kable(proportion_glasses, digits = 1, col.names = c("Age Bracket", "Proportion Wearing Glasses (%)"))
```
## C
```{r}
# Clean up gender
merged_data <- merged_data %>%
  rename(gender = RIAGENDR) %>%
  mutate(female = ifelse(gender == 2, 1, 0))
# Clean up race
merged_data <- merged_data %>%
  rename(race = RIDRETH1)
# Create labels for race categories (optional for visualization purposes)
race_labels <- c("Mexican American", "Other Hispanic", "Non-Hispanic White", "Non-Hispanic Black", "Multi-racial")
# Clean up PIR (poverty income ratio)
merged_data <- merged_data %>%
  rename(pir = INDFMPIR)
# Logistic regression models
# Model 1: age as predictor
model1 <- glm(glasses ~ age, data = merged_data, family = "binomial")
# Model 2: age, race, and gender as predictors
model2 <- glm(glasses ~ age + factor(race) + factor(female), data = merged_data, family = "binomial")
# Model 3: age, race, gender, and PIR as predictors
model3 <- glm(glasses ~ age + factor(race) + factor(female) + pir, data = merged_data, family = "binomial")

# Display the models' results with pseudo R-squared and AIC
stargazer::stargazer(model1, model2, model3, type = "text", 
                     apply.coef = exp,  # Show odds ratios
                     column.labels = c("Model 1", "Model 2", "Model 3"),
                     dep.var.labels = "Wears Glasses",
                     covariate.labels = c("Age", "Race: Other Hispanic", "Race: Non-Hispanic White", 
                                          "Race: Non-Hispanic Black", "Race: Multi-racial", "Female", "Poverty Income Ratio"),
                     add.lines = list(c("N", nobs(model1), nobs(model2), nobs(model3)),
                                      c("Pseudo R-squared", round(1 - model1$deviance/model1$null.deviance, 5),
                                                            round(1 - model2$deviance/model2$null.deviance, 5),
                                                            round(1 - model3$deviance/model3$null.deviance, 5)),
                                      c("AIC", round(AIC(model1), 2), round(AIC(model2), 2), round(AIC(model3), 2))))
```
## D
```{r}
# Load necessary libraries
if (!require(broom)) install.packages("broom")
if (!require(margins)) install.packages("margins")

library(broom)
library(margins)

# Fit the logistic regression model
model3 <- glm(glasses ~ age + factor(race) + factor(female) + pir, data = merged_data, family = "binomial")

# Get the odds ratios (exponentiate coefficients), standard errors, z-values, p-values, and confidence intervals
logit_results <- tidy(model3, conf.int = TRUE) %>%
  mutate(Odds_Ratio = exp(estimate), `2.5 %` = exp(conf.low), `97.5 %` = exp(conf.high)) %>%
  select(term, Odds_Ratio, std.error, statistic, p.value, `2.5 %`, `97.5 %`)

# Rename variables as per custom labels
logit_results <- logit_results %>%
  mutate(term = recode(term, 
                       `age` = "Age", 
                       `factor(race)2` = "Race: Other Hispanic",
                       `factor(race)3` = "Race: Non-Hispanic White",
                       `factor(race)4` = "Race: Non-Hispanic Black",
                       `factor(race)5` = "Race: Multi-racial",
                       `factor(female)1` = "Female",
                       `pir` = "Poverty Income Ratio"))

# Print the results with renamed columns and terms
print(logit_results)

# Calculate predictive margins for female
margins_female <- margins(model3, variables = "female")
summary(margins_female)

```
From the output, we know that the estimated odds ratio for females is 0.5967415, which is statistically significant, indicating that the odds of females wearing glasses of distance vision is statistically significantly lower than that of males. Also, from the tables, we observe that the proportion of female wearing glasses for distance vision is statistically lower than males.


# Problem 2
```{r}
# Load the dataset
library(DBI)
sakila <- dbConnect(RSQLite::SQLite(), "/Users/yining/Downloads/sakila_master.db")
```

## A
```{r}
# Do query
dbGetQuery(sakila, "
SELECT release_year, COUNT(*) AS movie_count
  FROM film
 GROUP BY release_year
 ORDER BY release_year ASC
 LIMIT 1;
")
```
The oldest movie is from 2006, and there were 1000 movies released that year.

## B
```{r}
# R approach
fc <- dbGetQuery(sakila, "SELECT * FROM film_category")
cat <- dbGetQuery(sakila, "SELECT * FROM category")
catcount <- table(fc$category_id)
mincat <- which.min(catcount)
c(cat$name[cat$category_id == mincat], catcount[mincat])
```

```{r}
# Single query approach
dbGetQuery(sakila, "
SELECT c.name, COUNT(c.category_id) AS count
  FROM category AS c
 RIGHT JOIN film_category AS fc ON fc.category_id = c.category_id
 GROUP BY c.category_id
 ORDER BY count ASC
 LIMIT 1
")
```
Music is the least genre, with 51 movies.

## C
```{r}
# Query the customer data from the 'customer' table in the sakila database
customer <- dbGetQuery(sakila, "SELECT * FROM customer")

# Query the address data from the 'address' table
address <- dbGetQuery(sakila, "SELECT * FROM address")

# Query the city data from the 'city' table
city <- dbGetQuery(sakila, "SELECT * FROM city")

# Query the country data from the 'country' table
country <- dbGetQuery(sakila, "SELECT * FROM country")

# Merge the customer data with the address data based on the 'address_id' column
# Keep all rows from 'customer' and only the matching rows from 'address' (left join)
merged1 <- merge(customer, address, by = "address_id", all.x = TRUE, all.y = FALSE)

# Merge the resulting dataset (merged1) with the city data based on the 'city_id' column
# Keep all rows from the previous merge (merged1) and only the matching rows from 'city'
merged2 <- merge(merged1, city, by = "city_id", all.x = TRUE, all.y = FALSE)

# Merge the resulting dataset (merged2) with the country data based on the 'country_id' column
# Keep all rows from the previous merge (merged2) and only the matching rows from 'country'
merged3 <- merge(merged2, country, by = "country_id", all.x = TRUE, all.y = FALSE)

# Create a frequency table showing the count of each country in the merged data
t <- table(merged3$country)

# Filter the table to display only countries that appear exactly 13 times
t[t == 13]

```

```{r}
# Find the city IDs corresponding to each customer by matching their address IDs in the address table
cities <- address$city_id[match(customer$address_id, address$address_id)]
# Find the country IDs corresponding to the city IDs found earlier
countries <- city$country_id[match(cities, city$city_id)]
# Create a frequency table of country names based on the country IDs found in the previous step.
tcountries <- table(country$country[match(countries, country$country_id)])
# Filter the frequency table to show only the countries that appear exactly 13 times
tcountries[tcountries == 13]

```

```{r}
# Single query approach
dbGetQuery(sakila, "
SELECT co.country, count(co.country) AS count
  FROM country AS co
 RIGHT JOIN
    (SELECT country_id
       FROM city AS ci
      RIGHT JOIN
         (SELECT city_id
            FROM customer AS c
            LEFT JOIN address AS a ON c.address_id = a.address_id
         ) AS ca ON ca.city_id = ci.city_id
    ) AS ccc ON ccc.country_id = co.country_id
 GROUP BY co.country
HAVING count == 13")
```

# Problem 3
```{r}
# Load the dataset
data <- read.csv("/Users/yining/Downloads/us-500.csv")
```
# A
```{r}
# calculate the proportion
length(data$email[grepl("com$", data$email)])/nrow(data)
```
# B
```{r}
# Split the email addresses into two parts: username (before '@') and domain (after '@')
emails <- strsplit(data$email, "@")

# Extract the username part (the part before the '@') from each email
username <- sapply(emails, "[[", 1)

# Check whether the username contains any non-alphanumeric characters
username_notalphanumeric <- grepl("[^a-zA-Z0-9]", username)

# Extract the domain part (the part after the '@') from each email
domains <- sapply(emails, "[[", 2)

# Remove the top-level domain (TLD) from the domain names
domains <- gsub("\\.[a-z]{3}", "", domains)

# Check whether the domain contains any non-alphanumeric characters
domain_notalphanumeric <- grepl("[^a-zA-Z0-9]", domains)

# Calculate the proportion of emails where either the username or the domain contains non-alphanumeric characters
mean(username_notalphanumeric | domain_notalphanumeric)

```
# C
```{r}
# We first check that all phone numbers have the same length of digits.
table(sapply(data$phone1, nchar))
# Extract the first three digits (area codes) from phone1 and phone2
phone1area <- substr(data$phone1, 1, 3)
phone2area <- substr(data$phone2, 1, 3)
# Combine the two area code vectors, count their occurrences, sort in decreasing order, and get the top 5
top5_area_codes <- sort(table(c(phone1area, phone2area)), decreasing = TRUE)[1:5]
# Display the result
top5_area_codes
```
# D
```{r}
# Filter out addresses that end with a number using regular expression
apartments <- data$address[grepl("[0-9]+$", data$address)]

# Split the filtered addresses by spaces and extract the last element (the apartment number)
numbers <- sapply(strsplit(apartments, " "), function(x) x[length(x)])

# Remove any "#" symbol from the numbers using 'gsub()'
numbers <- as.numeric(gsub("#", "", numbers))

# Create a histogram of the log-transformed apartment numbers
hist(log(numbers))

```
# E
```{r}
# Check Benfor's law
table(substr(numbers, 1, 1))
```
From the output, this is a normal distribution instead of a decreasing distribution by Benford's law, so the data wouldn't be pass as real.

Github Repo: https://github.com/romeor26/STATS-506